A geometric approach to linear Chabauty

### 2023-12-08 art.tex
Add defense files

{\Large\bf 
P.\ Spelier
}

\vspace{1em}

{\LARGE\bf 
A geometric approach to linear Chabauty\\
{\color{red} Over number fields}
}

\vspace{8em}

{\large\bf 
Master thesis
}

\vspace{1em}

@@

\vspace{8em}

{\large\bf
\begin{tabular}{ll}
Thesis supervisor: & prof. dr. S.J. Edixhoven\\
\end{tabular}
}
\vfill

\includegraphics{ulzegel_blauw}\\

{\large\bf 
Leiden University\\
Mathematical Institute\\
}

\end{center}
\thispagestyle{empty}
\newpage
\tableofcontents
\newpage

%-----------------------------------------------------------------------------------------------------------------------
\section{Introduction}
\label{section:intro}
A fundamental problem in mathematics is solving polynomial equations over the rationals, dating back to Diophantus. An important special case in algebraic geometry is that of curves. The behavior of the rational points of curves depends enormously on the \textit{genus} $g$ of the curve, a numerical invariant. For $g = 0$, there are either no or infinitely many solutions, and their behavior is well understood. For $g = 1$, we get either no point or an elliptic curve, whose set of rational points forms an often infinite group.

In this thesis, we look only at the case of a curve $C$ with genus $g > 1$. It turns out that there are always only finitely many rational points. This result was originally conjectured by Mordell in 1922 and finally proven by Faltings in 1983 in \cite{faltings}.

Before Faltings's theorem was proven, one of the major partial results was a theorem from Chabauty in 1941, phrased in terms of the rank $r$ of the group of rational points on the Jacobian $J$ of the curve, also called the Mordell-Weil rank, which is finite by the Mordell-Weil theorem. Chabauty proved, using $p$-adic methods, that if $r$ is strictly smaller than the genus $g$, then $C$ has only finitely many rational points. Loosely said, this proof and all Chabauty related theorems, rely on choosing a prime $p$ for which $C$ has good reduction, and intersecting $J(\Q)$ with $C(\Q_p)$ inside the bigger Lie group $J(\Q_p)$. By properties of $\Q_p$, the subgroup $J(\Q)$, up to torsion generated by $r$ elements, lies within a Lie subgroup of dimension at most $r$, and $C(\Q_p)$ is a $1$-dimensional $p$-adic manifold, and hence this intersection can be proven to be discrete and -- as $J(\Q_p)$ is compact -- finite.

@@

The purpose of this thesis is to make the work by Edixhoven and Lido more accessible by applying geometric Chabauty to the linear case, i.e. when working with $J(\Q)$ and $J(\Q_p)$ and assuming $r < g$.\todo{Add number fields in the introduction} In this thesis, we go into more detail on what is happening throughout the process, including both theoretical and practical components. This new method of applying Chabauty leads to, when working with single-digit $p$-adic precision, a conditional upper bound $|C(\Q)| \leq |C(\F_p)|$, as obtained in Proposition~\ref{prop:linchabauty}. This upper bound compares favorably with classical Coleman-Chabauty, but unlike classical Coleman-Chabauty, we cannot say in general when the conditions of these propositions hold. We develop heuristics that suggest that generally, trying multiple primes $p$ of good reduction will yield good results; these heuristics say that the conditional upper bound will hold for a set of primes of density $1$, and if $r < g-1$ the upper bound can even often (in a set of primes of density at least $e^{-1}$) be improved to prove there are at most $|C(\Q)|$ rational points, i.e., finding $C(\Q)$ exactly. The methods used to obtain Proposition~\ref{prop:linchabauty} are also amenable to higher precision calculations; indeed, in the case $r = g-1$ one expects to need either higher precision calculations or the Mordell-Weil sieve to calculate $|C(\Q)|$ exactly.

@@

\subsection{Overview}
We first present the context in which we will perform Chabauty. Let $K$ be a number field of degree $d$, and let $v: K \to \Z$ be a finite place. Let $\pi$ denote a uniformiser, $q$ the cardinality of the residue field and $p> 2$ its caracteristic\todo{Check when $p,\pi$ is big enough for both Pierre Parent and logarithm}\todo{Decide if we can allow ramification}. Let $f$ denote the residue field degree $[\F_q : \F_p]$ and $e$ the ramification $v(p)$. We use $O$ to denote the maximal order of $K$, and $O_{(\pi)}$ and $O_{\pi}$ for respectively the localisation and the completion with respect to $\pi$. This valuation also gives rise to an embedding of $K$ into an extension $K_{\pi}$ of $\Q_p$ of degree $ef$, with ramification degree $e$ and residue field degree $f$; then $O_\pi$ is simply the ring of integers of $K_{\pi}$.

Let $C_{K}/K$ be a curve (i.e., a proper, smooth, geometrically connected variety of dimension $1$) of genus $g \geq 2$ whose $K$-points we will attempt to find. assume we have a scheme $C$ over $O_{(\pi)}$, proper and smooth with generic fibre $C_{K}$; then we immediately have $C_K(K) = C(K) = C(O_{(\pi)})$ by the valuative criterion of properness. Let $J$ be the relative Jacobian of $C$ over $O_{(\pi)}$, i.e. with fibres $\Jac C_K$ and $\Jac C_{\F_q}$. We assume from now on that the Mordell-Weil rank of $J$ is $r \leq d(g-1)$. Assume we have a $K$-point $b$ in $C$, or equivalently a $O_{(\pi)}$-point. We view $C$ as a subscheme of $J$, using the map $Q \mapsto Q - b$ on points. Let $P \in C(\F_q)$ be a point such that $t := P - b \in J(\F_q)$ lies in the image of $J(O_{(\pi)})$.

\begin{defn}
Let $S$ be a scheme, $T \to U$ a morphism of schemes and $x : T \to S$ a $T$-point. We define $S(U)_x$ as the morphisms from $U$ to $S$ that, after precomposing with $T \to U$, give $x$.
\end{defn}
\begin{exmp}
If we have a proper variety $X$ over $O_{(\pi)}$, then $X(O_{(\pi)})$ is naturally in bijection with $X(K)$. The natural map $X(O_{(\pi)}) \to X(\F_q)$ reduces a point modulo $\pi$ and for $x \in X(\F_q)$, the set $X(O_{(\pi)})_x$ consists of the residue disc of $O_{(\pi)}$-points reducing to $x$.
\end{exmp}

Geometric Chabauty works by finding an upper bound for the cardinality of $C(O_{(\pi)})_P$. For this, we use the following diagram.
\[
\begin{tikzcd}
C(O_{(\pi)})_P \arrow[d] \arrow[r] & J(O_{(\pi)})_t \arrow[d] \\
C(O_\pi)_P \arrow[r]              & J(O_\pi)_t        
\end{tikzcd}
\]
where the two vertical maps are inclusions, and the two horizontal maps are subtraction of $b$. We will in fact compute upper bounds for the larger set $C(O_\pi)_P \cap \overline{J(O_{(\pi)})_t}$, where $\overline{J(O_{(\pi)})_t}$ is the closure of $J(O_{(\pi)})_t$ in $J(O_\pi)_t$.

In Section~\ref{section:smoothzppoints}, we treat the structure of $C(O_\pi)_P$ and $J(O_\pi)_t$; we will show that, after choosing parameters, they are in bijection with respectively $O_\pi$ and $O_\pi^g$. We also discuss the resulting map $O_\pi \to O_\pi^g$.{\color{red} and, identifying $O_\pi$ with $\Z_p^{ef}$ as topological groups, the map $\Z_p^{ef} \to \Z_p^{efg}$. This latter map is not necessarily well behaved if there is ramification; choosing a $\Z_p$-basis of $O_{\pi}$ appropriately, we will instead look at the well behaved restriction $\Z_p^{f} \to \Z_p^{fg}$.}

In Section~\ref{section:kappa} we further look at the map $J(O_{(\pi)})_t \to J(O_\pi)_t$. This is a translation of the group morphism $J(O_{(\pi)})_0 \to J(O_\pi)_0$ between kernels of reduction; it turns out that for our choice of $p$, the subgroup $J(O_{(\pi)})_0$ is free of rank $r$. We study the properties of the resulting map $\Z^r \to O_\pi^g$, using the theory of formal groups to determine the group structure on $O_\pi^g$ induced by the bijection $J(O_\pi)_0 \to O_\pi^g$. {\color{red} Importantly, we identify $O_\pi$ with $\Z_p^{ef}$ as topological groups; we also look at the resulting map $\Z^r \to \Z_p^{efg}$. Again, to control the behavior of this map modulo $p$, we consider the map $\Z^r \to \Z_p^{fg}$.}

In Section~\ref{section:intersection}, we put all of this information together, culminating in several methods to possibly compute upper bounds on $\left|C(O_\pi)_P \cap \overline{J(O_{(\pi)})_t}\right|$ with finite precision calculations if $r \leq f(g-1)$. We also perform a heuristic analysis to show that a simple calculation of linear algebra modulo $p$ is very likely to result in an upper bound of at most $|C(\F_q)|$ for $|C(O_{(\pi)})|$, and for $r < f(g-1)$ \todo{check whether this is correct, it should hold.} even in an upper bound of $|C_K(K)|$. Hence given only a curve $C_K/K$, the best possible choice of a prime of $K$ is an inert prime

The methods developed in the first four sections are not always guaranteed to prove finiteness of $C(O_{(\pi)})_P$. In Section~\ref{section:remarks} we treat several possible complications and improvements, referring to other work on Chabauty for further reading.

Next, in Section~\ref{section:explicit} we focus on how to perform the calculations happening in the Jacobian. Here, we use Makdisi's approach of representing a divisor by a subspace of a large Riemann-Roch space. We partially follow the article \cite{mascot18} by Mascot, and also give an original algorithm for explicitly computing the map $C(O_{\pi})_P \to J(O_{\pi})_t$ in Makdisi's representation.

We end our discussion of linear geometric Chabauty with an explicit example in Section~\ref{section:example}, a genus $2$ curve over the rationals whose Jacobian has Mordell-Weil rank $1$. We use both Magma and Pari/GP to do our calculations, and end up with a complete list of all rational points of the curve.

Finally, we provide an introduction to geometric quadratic Chabauty in Section~\ref{section:quadratic} (working only over the rationals). Here, we aim to introduce all the different objects relevant to geometric quadratic Chabauty; we explain why we have to work over $\Z$ instead of over $\Q$ in order for geometric quadratic Chabauty to make sense; and we explain some of the problems and solutions that working over $\Z$ gives rise to.

\subsection{Preliminaries}
We expect the readers of this thesis to be familiar with algebraic geometry, e.g. master students with a specialisation in algebraic geometry. The introduction to geometric quadratic Chabauty in Section~\ref{section:quadratic} contains some terms typically not treated in a master programme; while knowing them will help to put the material in context, this section is meant to be understandable for those that are not familiar with the terminology as well.

\subsection{Acknowledgements}
I would like to thank my supervisor Bas Edixhoven for his help --- I have learnt a great deal since starting on this thesis, and he helped me do so. I would also like to thank my father, for all the support he has shown me.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
% Say something about smooth schemes over Z_p, and their Z_p points
\section{Points on a smooth scheme over \texorpdfstring{$\Z_p$}{Z\_p}}
\label{section:smoothzppoints}
{\color{red} Most of this section should still hold in the general number field case; replacing $p$ by $\pi$ when it is an element of $A$ (i.e., a section of $X$) and by $q$ when it is the cardinality of the residue field. Only the remark on it being a torsor of the tangent space merits a checking+rewrite. This has been done.}
Let $X/\Z_p$ be a smooth scheme of relative dimension $d$, and let $x \in X(\F_p)$ be a point. Then we will show in this section that $X(\Z_p)_x$ is naturally in bijection with $\Z_p^d$. This bijection is given by choosing parameters at $x$; evaluating at $X(\Z_p)_x$ gives a bijection with $(p\Z_p)^d$, and then we divide by $p$. For putting up a nice framework to work in, we start with blowing up $X$ at $x$.

@@

\begin{lem}
There is a natural bijection \[X(\Z_p)_x \to \widetilde{X}_x^p(\Z_p).\]
\end{lem}
\begin{proof}
Note that by (I,2.4.4) of \cite{ega}, a $\Z_p$-point of a scheme $S$ is just an $\F_p$-point $s$ together with a local morphism $\\O_{S,s} \to \Z_p$. In our case, we find that $X(\Z_p)_x$ is naturally in bijection with $\Hom_{\text{local}}(A_x,\Z_p)$. As the maximal ideal of $A_x$ is generated by $p,t_1,\dots,t_d$, the morphism being local just means that the images of $t_1,\dots,t_d$ are divisible by $p$. That exactly gives those morphisms that extend to a morphism $A[t_1/p,\dots,t_d/p] \to \Z_p$, i.e. a $\Z_p$-point of $\widetilde{X}_x^p$. Hence we find the natural bijection.
\end{proof}

@@

\begin{rem}
\label{rem:torsortangentspace}
In actual calculations, we will be focusing on $X(S)_x$ where $S = O_\pi/\pi^2 O_\pi$. This set is a natural torsor of the tangent space $T_x X_{\F_q}$ of $X_{\F_q}$ at $x$, which we will describe here. We write $R$ for the local ring $\O_{X_{S},x}$ and $\m$ for its maximal ideal, and $\overline{R}$ and $\overline{\m}$ for their reductions modulo $\pi$. Then an element in $X(S)_x$ is a local morphism $R \to S$. As $R/\m = \F_q$, giving such a local morphism is equivalent to giving a map $\m \to \pi S$ respecting multiplication and sending $\pi$ to $\pi$. Such a map factors uniquely through $\m/\m^2 \to \pi S$. We see that the set $X(\Z/p^2\Z)_x$ is canonically in bijection with $\F_q$-linear maps $\m/\m^2 \to \pi S$ sending $\pi$ to $\pi$. Note that $\m/\m^2$ is a $(d+1)$-dimensional $\F_q$-vector space, and $\overline{\m}/\overline{\m}^2$ is a $d$-dimensional $\F_q$-vector space; the map $\m/\m^2 \to \overline{\m}/\overline{\m}^2$ is dividing out by $\pi$. Denoting ${}^\wedge$ to be $\Hom(\cdot,\F_q)$, the exact sequence
\[
0 \to \pi S \to \m/\m^2 \to \overline{\m}/\overline{\m}^2 \to 0
\]
can be dualised to
\[
0 \to (\overline{\m}/\overline{\m}^2)^\wedge \to (\m/\m^2)^\wedge \to (\pi S)^{\wedge} \to 0.
\]
As $X(S)_x$ is exactly the subset of $(\m/\m^2)^\wedge$ that gets mapped to the function $p \mapsto 1 \in (\pi S)^{\wedge}$, we see this is naturally a torsor under the tangent space $(\overline{\m}/\overline{\m}^2)^\wedge = T_{x} X_{\F_q}$.

@@

\begin{rem}
\label{rem:cartierlinear}
We look at the specific case of $X$ being of relative dimension $1$ over $\Z_p$. Then the $T_x X_{\F_p}$-torsor structure on $X(\Z/p^2\Z)_x$ translates into something more concrete. We can parametrise $X(\Z/p^2\Z)_x$ by $t = t_1 $ from $X(\Z/p^2\Z)_x \to p\Z/p^2\Z$; write $P_{\lambda}$ for the point with $t$-value $\lambda p$. Then as Cartier divisor, the point $P_{\lambda}$ is defined by $t - \lambda p \in \O_{X_{\Z/p^2\Z},x}$, so $P_{\lambda} + P_{\mu}$ is defined by $(t-\lambda p)(t-\mu p) = t^2 - (\lambda + \mu)tp$, which defines the same Cartier divisor as $P_{\lambda'} + P_{\mu'}$ if and only if $\lambda + \mu = \lambda' + \mu'$. In that case, as a Cartier divisor, $P_{\lambda} + P_{\mu}$ is in fact equal to $P_{\lambda'} + P_{\mu'}$.
\end{rem}

\subsection{From \texorpdfstring{$C(\Z_p)$}{C(Z\_p)} to \texorpdfstring{$J(\Z_p)$}{J(Z\_p)}}
\label{subsection:czptojzp}
Let $p,C$ and $J$ be as defined in the overview. We know that $C(\Z_p)_P$ is in bijection with $\Z_p$, again with the bijection given by evaluating a parameter and dividing by $p$. The resulting function $\Z_p \to \Z_p^g$ is linear modulo $p$, i.e. by using that $\Z\angle{z_1,\dots,z_g}$ is $p$-adically complete there are power series $f_1,\dots,f_{g-1} \in \Z_p\angle{z_1,\dots,z_g}$ such that the image of $C(\Z_p)_P$ is exactly given by $Z(f_1,\dots,f_{g-1})$, and all $f_i$ are linear modulo $p$ as a consequence of Remark \ref{rem:cartierlinear}. Another way to think of this, is as $C(\Z/p^2\Z)_P$ being an affine line inside $J(\Z/p^2\Z)_t$. That $C(\Z/p^2\Z)_P$ is an affine line inside $J(\Z/p^2\Z)_t$ can also be seen more easily: this map can be identified with the tangent map, using the structure of $X(\Z/p^2\Z)_x$ as a torsor over the tangent space from Remark~\ref{rem:cartierlinear}.

@@

@@

Of note is that this upper bound for $|C(\Z_{(p)})|$ does not even depend on the genus, and behaves agreeably when compared to the upper bound in Theorem 5.3 of \cite{poonen12}, which tells us that $|C(\Z_{(p)})| \leq |C(\F_p)| + 2g-2$ under the much lighter condition of $r < g, p > 2g$ where $p$ is still a prime of good reduction. To summarise: in Proposition~\ref{prop:linchabauty} we sacrifice certainty for being able to prove much sharper bounds. However, the loss of certainty is not too great; a simple heuristic analysis shows that the conditions are quite likely to be satisfied, as we shall see in the following lemma and proposition.

@@

\section{Quadratic Chabauty}
\label{section:quadratic}
In this section, we attempt to give an introduction to the article~\cite{edixhoven20} aimed to introduce all the players in quadratic Chabauty at a slightly slower pace. We will omit all proofs. For further reading, see for example \cite{hashimoto20}, an in-progress cartoon guide to the article by Edixhoven and Lido.

>>>>

{\Large\bf 
P.\ Spelier, M.Sc\\
prof. dr. S.J. Edixhoven
}

{\sc 
Leiden University, Mathematical Institute\\
}

{\LARGE\bf 
A geometric approach to linear Chabauty
} 
\vspace{2em}

@@

\end{center}

\listoftodos
\todo{Bas, zou je wellicht een bewijs kunnen schrijven dat onder de voorwaarde $r < g$ (ik weet niet of er nog andere nodig zijn) de verzameling gedefinieerd in Theorem~\ref{thm:final} altijd eindig is? Dat lijkt me fijn, en misschien geeft het zelfs een algoritme dat voor elke kromme een bovengrens kan geven?}
%-----------------------------------------------------------------------------------------------------------------------

\todo{First try of abstract}
Given a curve of genus at least $2$, it is known by Faltings's Theorem (1983) that the curve has only finitely many rational points. Finding the set of rational points can still be a difficult problem. Chabauty (1941) already proved the special case of Faltings's Theorem where the Mordell-Weil rank of the Jacobian is less than the genus, by intersecting the $p$-adic points of the curve with the Mordell-Weil group inside the $p$-adic points of the Jacobian, and Coleman (1985) gave a way to make this effective, by introducing Coleman integrals. This article aims to make Chabauty's method effective in a different, purely geometric way. An important part of this is working with finite precision; we will in particular focus on working modulo $p$, and will develop several heuristics, and also calculate the complexity of our algorithms.

\tableofcontents

\section{Old introduction} \todo{Put a bit of the introduction, and some of the references/results in the new introduction}
\label{section:intro}
A fundamental problem in mathematics is solving polynomial equations over the rationals, dating back to Diophantus. An important special case in algebraic geometry is that of curves. The behavior of the rational points of curves depends enormously on the \textit{genus} $g$ of the curve, a numerical invariant. For $g = 0$, there are either no or infinitely many solutions, and their behavior is well understood. For $g = 1$, we get either no point or an elliptic curve, whose set of rational points forms a finitely generated abelian group.

In this article, we look only at the case of a curve $C$ with genus $g > 1$. It turns out that there are always only finitely many rational points. This result was originally conjectured by Mordell in 1922 and finally proven by Faltings in 1983 in \cite{faltings}.

Before Faltings's theorem was proven, one of the major partial results was a theorem from Chabauty in 1941, phrased in terms of the rank $r$ of the group of rational points on the Jacobian $J$ of the curve, also called the Mordell-Weil rank; this rank is finite by the Mordell-Weil theorem. Chabauty proved, using $p$-adic methods, that if $r$ is strictly smaller than the genus $g$, then $C$ has only finitely many rational points. Loosely said, this proof and all Chabauty-related theorems, rely on choosing a prime $p$ for which $C$ has good reduction, and intersecting $J(\Q)$ with $C(\Q_p)$ inside the bigger Lie group $J(\Q_p)$. By properties of $\Q_p$, the subgroup $J(\Q)$, up to torsion generated by $r$ elements, lies within a Lie subgroup of dimension at most $r$, and $C(\Q_p)$ is a $1$-dimensional $p$-adic manifold, and hence this intersection can be proven to be discrete and -- as $J(\Q_p)$ is compact -- finite.

@@

The purpose of this article is to make the work by Edixhoven and Lido more accessible by applying geometric Chabauty to the linear case, i.e. when working with $J(\Q)$ and $J(\Q_p)$ and assuming $r < g$. In this article, we go into more detail on what is happening throughout the process, including both theoretical and practical components. This new method of applying Chabauty leads to, when working with single-digit $p$-adic precision, a conditional upper bound $|C(\Q)| \leq |C(\F_p)|$, as obtained in Proposition~\ref{prop:linchabauty}. This upper bound compares favorably with classical Coleman-Chabauty, but unlike classical Coleman-Chabauty, we cannot say in general when the conditions of these propositions hold. We develop heuristics that suggest that generally, trying multiple primes $p$ of good reduction will yield good results; these heuristics say that the conditional upper bound will hold for a set of primes of density $1$, and if $r < g-1$ the upper bound can even often (in a set of primes of density at least $e^{-1}$) be improved to prove there are at most $|C(\Q)|$ rational points, i.e., finding $C(\Q)$ exactly. The methods used to obtain Proposition~\ref{prop:linchabauty} are also amenable to higher precision calculations; indeed, in the case $r = g-1$ one expects to need either higher precision calculations or the Mordell-Weil sieve to calculate $|C(\Q)|$ exactly.

@@

\section{Introduction}\todo{Maybe give a bit more context (Faltings, Chabauty, Coleman, McCallum-Poonen, Edixhoven-Lido)? Not sure what I want to put in and what not.}\todo{Verwijzen naar dat Sachi en ik eraan werken om de connectie tussen Coleman en geometric duidelijker te maken.}
We first present the context in which we will perform Chabauty. Let $C_{\Q}/\Q$ be a curve (i.e., a proper, smooth, geometrically connected variety of dimension~$1$) of genus $g \geq 2$ whose $\Q$-points we will attempt to find. Let $p>2$ be a prime, and assume we have a scheme $C$ over $\Z_{(p)}$, proper and smooth with generic fibre $C_{\Q}$; then we immediately have $C_\Q(\Q) = C(\Q) = C(\Z_{(p)})$ by the valuative criterion of properness. Let $J$ be the relative Jacobian of $C$ over $\Z_{(p)}$, i.e. with fibres $\Jac C_\Q$ and $\Jac C_{\F_p}$. We assume from now on that the Mordell-Weil rank of $J$ is $r < g$. Assume we have a $\Q$-point $b$ in $C$, or equivalently a $\Z_{(p)}$-point. We view $C$ as a subscheme of $J$, using the map $Q \mapsto Q - b$ on points. Let $P \in C(\F_p)$ be a point such that $t := P - b \in J(\F_p)$ lies in the image of $J(\Z_{(p)})$.

\begin{defn}
Let $S$ be a scheme, $T \to U$ a morphism of schemes and $x : T \to S$ a $T$-point. We define $S(U)_x$ as the morphisms from $U$ to $S$ that, after precomposing with $T \to U$, give $x$.
\end{defn}
\begin{exmp}
If we have a proper variety $X$ over $\Z_{(p)}$, then $X(\Z_{(p)})$ is naturally in bijection with $X(\Q)$. The natural map $X(\Z_{(p)}) \to X(\F_p)$ reduces a point modulo $p$ and for $x \in X(\F_p)$, the set $X(\Z_{(p)})_x$ consists of the residue disc of $\Z_{(p)}$-points reducing to $x$.
\end{exmp}

Geometric Chabauty works by finding an upper bound for the cardinality of $C(\Z_{(p)})_P$. For this, we use the following diagram.
\[
\begin{tikzcd}
C(\Z_{(p)})_P \arrow[d] \arrow[r] & J(\Z_{(p)})_t \arrow[d] \\
C(\Z_p)_P \arrow[r]              & J(\Z_p)_t        
\end{tikzcd}
\]
where the two vertical maps are inclusions, and the two horizontal maps are subtraction of $b$. We will in fact compute upper bounds for the larger set $C(\Z_p)_P \cap \overline{J(\Z_{(p)})_t}$, where $\overline{J(\Z_{(p)})_t}$ is the closure of $J(\Z_{(p)})_t$ in $J(\Z_p)_t$.

In goemetric Chabauty, in contrast to Coleman-Chabauty (as explained in \cite{poonen12}), we give equations for the curve inside the Jacobian, and pull those back to the Mordell-Weil group. The main result is Theorem~\ref{thm:final}, which in the case of working modulo $p$ specialises to Proposition~\ref{prop:linchabauty}. This proposition gives a conditional upper bound on $|C(\Q)|$ of at most $|C(\F_p)|$. This behaves comparably with Theorem~5.3 of \cite{poonen12}, which gives an upper bound of $|C(\F_p)| + 2g-2$, albeit under less strict conditions.

It turns out that, when working modulo $p$, the situation becomes linear algebra; we give some heuristics about how often the conditions of Proposition~\ref{prop:linchabauty} are satisfied, and what the upper bound will be in Proposition~\ref{prop:probwork} and Remark~\ref{rem:probwork}. In particular, we expect that our method will give an upper bound on $|C(\Q)|$ for all $p$ in a set of density $1$, and for $r \leq g-2$ will even lead to the optimal upper bound of $|C(\Q)|$ in a set of density at least $e^{-1}$ (with both results having refinements for even smaller $r$).

In this article, we also focus on explicit algorithms and computations. Assuming all relevant computations in the Mordell-Weil group have been done (finding the rank, a set of generators, and other data), in Section~\ref{section:explicit} we give an $O((p+g\sqrt{p})rg^{4})$ algorithm for executing the calculations needed in Proposition~\ref{prop:linchabauty}, and we implement and run this in Pari/GP.

\subsection{Overview}
In Section~\ref{section:smoothzppoints}, we treat the structure of $C(\Z_p)_P$ and $J(\Z_p)_t$; we will show that, after choosing parameters, they are in bijection with respectively $\Z_p$ and $\Z_p^g$. We also discuss the resulting map $\Z_p \to \Z_p^g$.

In Section~\ref{section:kappa} we further look at the map $J(\Z_{(p)})_t \to J(\Z_p)_t$. This is a translation of the group morphism $J(\Z_{(p)})_0 \to J(\Z_p)_0$ between kernels of reduction; it turns out that for our choice of $p$, the subgroup $J(\Z_{(p)})_0$ is free of rank $r$. We study the properties of the resulting map $\Z^r \to \Z_p^g$, using the theory of formal groups to determine the group structure on $\Z_p^g$ induced by the bijection $J(\Z_p)_0 \to \Z_p^g$.

In Section~\ref{section:intersection}, we put all of this information together, culminating in several methods to possibly compute upper bounds on $\left|C(\Z_p)_P \cap \overline{J(\Z_{(p)})_t}\right|$ with finite precision calculations. We also perform a heuristic analysis to show that a simple calculation of linear algebra modulo $p$ is very likely to result in an upper bound of at most $|C(\F_p)|$ for $|C(\Z_{(p)})|$, and for $r < g-1$ even in an upper bound of $|C(\Q)|$.

The methods developed in the first four sections are not always guaranteed to prove finiteness of $C(\Z_{(p)})_P$. In Section~\ref{section:remarks} we treat several possible complications and improvements, referring to other work on Chabauty for further reading.

Next, in Section~\ref{section:explicit} we focus on how to perform the calculations happening in the Jacobian. Here, we use Makdisi's approach of representing a divisor by a subspace of a large Riemann-Roch space. We partially follow the article \cite{mascot18} by Mascot, and also give an algorithm for explicitly computing the map $C(\Z_p)_P \to J(\Z_p)_t$ in Makdisi's representation.

We end our discussion of linear geometric Chabauty with an explicit example in Section~\ref{section:example}, a genus $2$ curve whose Jacobian has Mordell-Weil rank $1$. We use both Magma and Pari/GP to do our calculations, and end up with a complete list of all rational points of the curve.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
% Say something about smooth schemes over Z_p, and their Z_p points
\section{Points on a smooth scheme over \texorpdfstring{$\Z_p$}{Z\_p}}
\label{section:smoothzppoints}
Let $X/\Z_p$ be a smooth scheme of relative dimension $d$, and let $x \in X(\F_p)$ be a point. Then we will show in this section that $X(\Z_p)_x$ is, up to a single choice, naturally in bijection with $\Z_p^d$. This bijection is given by choosing parameters at $x$; evaluating at $X(\Z_p)_x$ gives a bijection with $(p\Z_p)^d$, and then we divide by $p$. For putting up a nice framework to work in, we start with blowing up $X$ at $x$.

@@

\begin{lem}
There is a natural bijection \[X(\Z_p)_x \to \widetilde{X}_x^p(\Z_p).\]
\end{lem}
\begin{proof}
Note that by (I,2.4.4) of \cite{ega}, a $\Z_p$-point of a scheme $S$ is just an $\F_p$-point $s$ together with a local morphism $\O_{S,s} \to \Z_p$. In our case, we find that $X(\Z_p)_x$ is naturally in bijection with $\Hom_{\text{local}}(A_x,\Z_p)$. As the maximal ideal of $A_x$ is generated by $p,t_1,\dots,t_d$, the morphism being local just means that the images of $t_1,\dots,t_d$ are divisible by $p$. That exactly gives those morphisms that extend to a morphism $A[t_1/p,\dots,t_d/p] \to \Z_p$, i.e. a $\Z_p$-point of $\widetilde{X}_x^p$. Hence we find the natural bijection.
\end{proof}

@@

\begin{rem}
\label{rem:torsortangentspace}
In actual calculations, we will be focusing on $X(\Z/p^2\Z)_x$. This set is a natural torsor of the tangent space $T_x X_{\F_p}$ of $X_{\F_p}$ at $x$, which we will describe here. We write $R$ for the local ring $\O_{X_{\Z/p^2\Z},x}$ and $\m$ for its maximal ideal, and $\overline{R}$ and $\overline{\m}$ for their reductions modulo $p$. Then an element in $X(\Z/p^2\Z)_x$ is a local morphism $R \to \Z/p^2\Z$. As $R/\m = \F_p$, giving such a local morphism is equivalent to giving a map $\m \to p\Z/p^2\Z$ respecting multiplication and sending $p$ to $p$. Such a map factors uniquely through $\m/\m^2 \to p\Z/p^2\Z$. We see that the set $X(\Z/p^2\Z)_x$ is canonically in bijection with $\F_p$-linear maps $\m/\m^2 \to p\Z/p^2\Z$ sending $p$ to $p$. Note that $\m/\m^2$ is a $(d+1)$-dimensional $\F_p$-vector space, and $\overline{\m}/\overline{\m}^2$ is a $d$-dimensional $\F_p$-vector space; the map $\m/\m^2 \to \overline{\m}/\overline{\m}^2$ is dividing out by $p$. Denoting ${}^\wedge$ to be $\Hom(\cdot,\F_p)$, the exact sequence
\[
0 \to p\Z/p^2\Z \to \m/\m^2 \to \overline{\m}/\overline{\m}^2 \to 0
\]
can be dualised to
\[
0 \to (\overline{\m}/\overline{\m}^2)^\wedge \to (\m/\m^2)^\wedge \to (p\Z/p^2\Z)^{\wedge} \to 0.
\]
As $X(\Z/p^2\Z)_x$ is exactly the subset of $(\m/\m^2)^\wedge$ that gets mapped to the function $p \mapsto 1 \in (p\Z/p^2\Z)^{\wedge}$, we see this is naturally a torsor under the tangent space $(\overline{\m}/\overline{\m}^2)^\wedge = T_{x} X_{\F_p}$.

@@

\begin{rem}
\label{rem:cartierlinear}
We look at the specific case of $X$ being of relative dimension~$1$ over $\Z_p$. Then the $T_x X_{\F_p}$-torsor structure on $X(\Z/p^2\Z)_x$ translates into something more concrete. We can parametrise $X(\Z/p^2\Z)_x$ by $t = t_1 $ from $X(\Z/p^2\Z)_x \to p\Z/p^2\Z$; write $P_{\lambda}$ for the point with $t$-value $\lambda p$. Then as a Cartier divisor, the point $P_{\lambda}$ is defined by $t - \lambda p \in \O_{X_{\Z/p^2\Z},x}$, so $P_{\lambda} + P_{\mu}$ is defined by $(t-\lambda p)(t-\mu p) = t^2 - (\lambda + \mu)tp$, which defines the same Cartier divisor as $P_{\lambda'} + P_{\mu'}$ if and only if $\lambda + \mu = \lambda' + \mu'$. In that case, as a Cartier divisor, $P_{\lambda} + P_{\mu}$ is in fact equal to $P_{\lambda'} + P_{\mu'}$.
\end{rem}

\subsection{From \texorpdfstring{$C(\Z_p)$}{C(Z\_p)} to \texorpdfstring{$J(\Z_p)$}{J(Z\_p)}}
\label{subsection:czptojzp}
Let $p$, $C$ and $J$ be as defined in the overview. We know that $C(\Z_p)_P$ is in bijection with $\Z_p$, again with the bijection given by evaluating a parameter and dividing by $p$. The resulting function $\Z_p \to \Z_p^g$ is linear modulo $p$, i.e. by using that $\Z\angle{z_1,\dots,z_g}$ is $p$-adically complete there are power series $f_1,\dots,f_{g-1} \in \Z_p\angle{z_1,\dots,z_g}$ such that the image of $C(\Z_p)_P$ is exactly given by $Z(f_1,\dots,f_{g-1})$, and all $f_i$ are linear modulo $p$ as a consequence of Remark \ref{rem:cartierlinear}. Another way to think of this, is as $C(\Z/p^2\Z)_P$ being an affine line inside $J(\Z/p^2\Z)_t$. That $C(\Z/p^2\Z)_P$ is an affine line inside $J(\Z/p^2\Z)_t$ can also be seen more easily: this map can be identified with the tangent map, using the structure of $X(\Z/p^2\Z)_x$ as a torsor over the tangent space from Remark~\ref{rem:cartierlinear}.

@@

\begin{exmp}\todo{If I can't find a reference, at least remove this examples.}

@@

Of note is that this upper bound for $|C(\Z_{(p)})|$ behaves agreeably when compared to the upper bound in Theorem 5.3 of \cite{poonen12}, which tells us that $|C(\Z_{(p)})| \leq |C(\F_p)| + 2g-2$ under the much lighter condition of $r < g, p > 2g$ where $p$ is still a prime of good reduction. To summarise: in Proposition~\ref{prop:linchabauty} we sacrifice certainty for being able to prove much sharper bounds. However, the loss of certainty is not too great; a simple heuristic analysis shows that the conditions are quite likely to be satisfied, as we shall see in the following lemma and proposition.

@@
